<<<<<<< HEAD
---
title: "Maxquant proteinGroups.txt Quick summary"
author: suggestions please to "ningzhibinn@gmail.com"
date: "`r Sys.time()`"
output:
  html_document:
    fig_width: 10
    fig_caption: TRUE
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: TRUE
      smooth_scroll: TRUE
    number_sections: true
params:  
  input_datatable: !r mtcars ### note that the value is the real value, here supposed to be a read in data.table for one click function.  
---

<style type="text/css">
.main-container {
  max-width: 1400px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{r setup, include=FALSE}
# enviroment setup
knitr::opts_chunk$set(echo = FALSE,warning = FALSE, message = FALSE, cache = TRUE)

library(tidyverse)
library(ggplot2)
library(d3heatmap)
library(plotly)
library(gplots)
#library(corrplot)
#library(DT) # for interactive data table display 

# local usage test and debug
# rmarkdown::render("MQ_report_summary_ocpu.Rmd", params = list(summary_file_tbl =  your_readin_tbl))
# on local machine, your_readin_tbl is a data table, while on ocpu, your_readin_tbl is a json formatted table

# version control
# 20190807



```



```{r function_needed, recho=FALSE}

# proteinGroups <- proteinGroups.txt

tidy_proteingroups <- function(proteinGroups){ 

  # extract the primary protein ID,
  protein.ids_split <- strsplit(as.vector(proteinGroups$"Protein IDs"), ";") # this is a list of list of split names
  protein_primary_ids <- unlist(lapply(protein.ids_split, function(x) x[1])) # only keep the first one
  #rownames(proteinGroups) <- protein_primary_ids # rename the rownames of the matrix
  
  # do the row wise filtering
  index_contaminant <- grep("\\+", proteinGroups$`Potential contaminant`) # note that + is a special character
  index_reverse <- grep("\\+", proteinGroups$Reverse)
  index_to_remove <- c(index_contaminant,index_reverse)
  
  proteinGroups <- proteinGroups[-index_to_remove,] # filtered table
  protein_primary_ids <- protein_primary_ids[-index_to_remove] # filtered ids

  n_contaminant <- length(index_contaminant)
  n_reversed <- length(index_reverse)

  
  # extra the intensity column matrix
  if(any(grepl("LFQ intensity ", colnames(proteinGroups)))){ # if there are LFQ intensity columns, take out the LFQ columns
    intensity_columns <- proteinGroups[,grep("LFQ intensity ", colnames(proteinGroups))]
    colnames(intensity_columns)<-gsub("LFQ intensity ", "", colnames(intensity_columns))
  }else if(any(grepl("Intensity ", colnames(proteinGroups)))) { # otherwise take out intensity column
    intensity_columns <-   proteinGroups[,grep("Intensity ", colnames(proteinGroups))]
    colnames(intensity_columns)<-gsub("Intensity ", "", colnames(intensity_columns))
  }else{ # this is for the special case that no experiment desgin setup, only one column of intensity
    intensity_columns <-   proteinGroups[,grep("Intensity", colnames(proteinGroups)),drop =  FALSE]
  } 
  
  return(list("intensity_matrix" = intensity_columns,
              "n_contaminant" = n_contaminant,
              "n_reversed" = n_reversed,
              "n_unique_peptides" = proteinGroups$"Unique peptides",
              "score" = proteinGroups$Score,
              "protein_primary_ids" =protein_primary_ids
              
              ))
  
}



```



```{r file_summary, echo=FALSE,fig.width= 15,fig.height=10}


# start from the data table for parametrized version
proteingroups_tidyed <- tidy_proteingroups(params$input_datatable)


# This 2 lines is only for debugging
#proteinGroups.txt  <- read.delim("proteinGroups4.txt", header = TRUE,check.names = FALSE, stringsAsFactors = FALSE)
#proteingroups_tidyed<- tidy_proteingroups(proteinGroups.txt)


df_intensity <- proteingroups_tidyed$intensity_matrix


sparsity <- rowSums(df_intensity > 0) # here sparsity is number of present values

index_all_na_rows <- which(sparsity == 0)

df_intensity <- df_intensity[-index_all_na_rows,,drop = FALSE]

protein_primary_ids <- proteingroups_tidyed$protein_primary_ids[-index_all_na_rows]



```

# Intro

**This report provides some basic description of the proteingroups identificaiton from maxquant database search. **
**The report is based on the proteinGroups.txt, without defined experimental design/grouping information**
**Users can use this to quickly check the overal quality of the experiment**


# Take home figures



*  **Number of contaminant: ** `r proteingroups_tidyed$n_contaminant`

*  **Number of reversed: ** `r proteingroups_tidyed$n_reversed`

*  **Number of qualified proteingroups: ** `r nrow(proteingroups_tidyed$intensity_matrix)`

*  **Number of quailfied proteingroups without intensity information:  ** `r length(index_all_na_rows)`

*  **Number of experiment:  ** `r ncol(df_intensity)`

*  **All experiments:  ** `r colnames(df_intensity)`




# Unique peptide number


```{r echo=FALSE,fig.width= 8,fig.height=6}

peptide_summary <- as.data.frame(table(proteingroups_tidyed$n_unique_peptides))
colnames(peptide_summary) = c("Unique_Peptides", "Freq")


p<- ggplot2::ggplot(data = peptide_summary)+
  geom_col(aes(x = Unique_Peptides,y = Freq))+
  theme_bw()
plotly::ggplotly(p)




```


# Protein Score distribution


```{r echo=FALSE,fig.width= 8,fig.height=6}

df_score <- data.frame(Score  = proteingroups_tidyed$score)


p<- ggplot2::ggplot(data=df_score, aes(Score)) +
  geom_histogram(aes(y=..density..),color= "black",fill="white")+
  geom_density(alpha=.2, fill="#FF6666")+
  theme_bw()

plotly::ggplotly(p)

```


# Intensity distribution

Columns starting with "LFQ_intensity_" will be selected if exist (meaning LFQ option was checked for label free quantification, which is deliberately processed by Maxquant already), otherwise Intensity_ columns will be used instead for protein experession.





```{r echo=FALSE,fig.width= 8,fig.height=6}

df_sparsity <- as.data.frame(table(sparsity))


p<- ggplot2::ggplot(data = df_sparsity)+
  geom_col(aes(x = sparsity,y = Freq))+
  xlab("Presence in the Protein Intensity Matrix")+
  theme_bw()
plotly::ggplotly(p)



```


```{r echo=FALSE,fig.width= 8,fig.height=6}

#note: in this log10 intenisty matrix, 0 is converted into infinity, therefore will not show in the box plot, and does not affect the distribution
df_intensity_log10 <-log10(df_intensity)

data_matrix_log10_melt<-reshape2::melt(as.matrix(df_intensity_log10))

colnames(data_matrix_log10_melt) <- c("Proteins", "Samples", "Log10(Inensity)")

p<-ggplot(data_matrix_log10_melt)+
    geom_boxplot(aes(x = Samples, y = `Log10(Inensity)`)) +
    theme_bw()+
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
  

plotly::ggplotly(p)


```


`r if(ncol(df_intensity) > 1){ "# Overall Expression Profile   ## Heatmap"} `



```{r echo=FALSE,fig.width= 8,fig.height=6}

# only if there is experiment desgin and  number of columns in protein experssion value is more than 1, 
if(ncol(df_intensity) > 1){ 
  df_intensity_log10 <-log10(df_intensity+1)
  rownames(df_intensity_log10) <- protein_primary_ids
  
  d3heatmap(df_intensity_log10,show_grid = FALSE, color = "OrRd")


  
}


```



`r if(ncol(df_intensity) > 1){ "## PCA"}`


```{r echo=FALSE,fig.width= 8,fig.height=6}

if(ncol(df_intensity) > 1){ 
  
  PCA_result <- prcomp(t(df_intensity_log10))
  loading <- as.data.frame(PCA_result$x)
  loading$Sample.Name <- rownames(loading)
  plot_ly(loading, x = ~PC1, y = ~PC2, z = ~PC3) %>%
  add_markers() %>%
  add_text(text = row.names(loading))

}


```

`r if(ncol(df_intensity) > 1){ "**Scree plot shows the performance of the PCA analysis, the more the top procomponets accounts, the better separation**"}`


```{r echo=FALSE,fig.width= 8,fig.height=6}

  # screen plot
if(ncol(df_intensity) > 1){ 

  sd <- PCA_result$sde
  var <- sd^2
  var.percent <- var/sum(var) * 100
  
  PCs <- paste("PC", 1:length(var.percent))
  
  
  df_scree <- data.frame(PC  = factor(PCs, levels = PCs), ratio =var.percent)
  
  
  plot_ly(data = df_scree,x = ~PC, y = ~ratio, type = "bar") %>% 
    layout(title = "Scree Plot of Principle components", xaxis = list(title = 'Principle Component'), yaxis = list(title = 'Variance(%)'))
  

  }


```



# Download the neat table

**The table is clean proteingroups expression table, with reversed and contaminant removed. Only the first protein id was used to represent the proteingroups.**
**The values are the LFQ(with label free quantification turned on) or raw protein intensity from Maxquant output. **
**You can download/export and start from this table for downstream analysis using the (i)Metalab faminly apps. **


```{r echo=FALSE,fig.width= 8,fig.height=6}


df_intensity <- as.data.frame(df_intensity)
rownames(df_intensity) <- protein_primary_ids


DT::datatable(df_intensity, extensions = 'Buttons',options = list(dom = "Blfrtip",scrollX=TRUE, buttons =       list('copy', 'print', list(
        extend = 'collection',
        buttons = c('csv', 'excel', 'pdf'),
        text = 'Download'
      ))))




```



```{r}
# library(ggforce)
# 
# a1 <-100
# a2 <- 500
# a3 <- 1000
# 
# r1 <- sqrt(a1)
# r2 <- sqrt(a2)
# r3 <-sqrt(a3)
# 
# 
# 
# 
# circles <- data.frame(
#   x0 = rep(0,3),
#   y0 = sort(c(r1,r2,r3),decreasing =TRUE),
#   r = sort(c(r1,r2,r3),decreasing =TRUE)
# )
# 
# ggplot(data = circles) +
#   geom_circle( aes(x0 = x0, y0 = y0, r = r, alpha = rev(r)), color = "orange", fill= "orange",show.legend = c("a","b","c") )+
#   coord_fixed()+
#   theme_void()




```





=======
---
title: "Maxquant proteinGroups.txt Quick summary"
author: suggestions please to "ningzhibinn@gmail.com"
date: "`r Sys.time()`"
output:
  html_document:
    fig_width: 10
    fig_caption: TRUE
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: TRUE
      smooth_scroll: TRUE
    number_sections: true
params:  
  input_datatable: !r mtcars ### note that the value is the real value, here supposed to be a read in data.table for one click function.  
---

<style type="text/css">
.main-container {
  max-width: 1400px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{r setup, include=FALSE}
# enviroment setup
knitr::opts_chunk$set(echo = FALSE,warning = FALSE, message = FALSE, cache = TRUE)

library(tidyverse)
library(ggplot2)
library(d3heatmap)
library(plotly)
library(gplots)
#library(corrplot)
#library(DT) # for interactive data table display 

# local usage test and debug
# rmarkdown::render("MQ_report_summary_ocpu.Rmd", params = list(summary_file_tbl =  your_readin_tbl))
# on local machine, your_readin_tbl is a data table, while on ocpu, your_readin_tbl is a json formatted table

# version control
# 20190807



```



```{r function_needed, recho=FALSE}

# proteinGroups <- proteinGroups.txt

tidy_proteingroups <- function(proteinGroups){ 

  # extract the primary protein ID,
  protein.ids_split <- strsplit(as.vector(proteinGroups$"Protein IDs"), ";") # this is a list of list of split names
  protein_primary_ids <- unlist(lapply(protein.ids_split, function(x) x[1])) # only keep the first one
  #rownames(proteinGroups) <- protein_primary_ids # rename the rownames of the matrix
  
  # do the row wise filtering
  index_contaminant <- grep("\\+", proteinGroups$`Potential contaminant`) # note that + is a special character
  index_reverse <- grep("\\+", proteinGroups$Reverse)
  index_to_remove <- c(index_contaminant,index_reverse)
  
  proteinGroups <- proteinGroups[-index_to_remove,] # filtered table
  protein_primary_ids <- protein_primary_ids[-index_to_remove] # filtered ids

  n_contaminant <- length(index_contaminant)
  n_reversed <- length(index_reverse)

  
  # extra the intensity column matrix
  if(any(grepl("LFQ intensity ", colnames(proteinGroups)))){ # if there are LFQ intensity columns, take out the LFQ columns
    intensity_columns <- proteinGroups[,grep("LFQ intensity ", colnames(proteinGroups))]
    colnames(intensity_columns)<-gsub("LFQ intensity ", "", colnames(intensity_columns))
  }else if(any(grepl("Intensity ", colnames(proteinGroups)))) { # otherwise take out intensity column
    intensity_columns <-   proteinGroups[,grep("Intensity ", colnames(proteinGroups))]
    colnames(intensity_columns)<-gsub("Intensity ", "", colnames(intensity_columns))
  }else{ # this is for the special case that no experiment desgin setup, only one column of intensity
    intensity_columns <-   proteinGroups[,grep("Intensity", colnames(proteinGroups)),drop =  FALSE]
  } 
  
  return(list("intensity_matrix" = intensity_columns,
              "n_contaminant" = n_contaminant,
              "n_reversed" = n_reversed,
              "n_unique_peptides" = proteinGroups$"Unique peptides",
              "score" = proteinGroups$Score,
              "protein_primary_ids" =protein_primary_ids
              
              ))
  
}



```



```{r file_summary, echo=FALSE,fig.width= 15,fig.height=10}


# start from the data table for parametrized version
proteingroups_tidyed <- tidy_proteingroups(params$input_datatable)


# This 2 lines is only for debugging
#proteinGroups.txt  <- read.delim("proteinGroups4.txt", header = TRUE,check.names = FALSE, stringsAsFactors = FALSE)
#proteingroups_tidyed<- tidy_proteingroups(proteinGroups.txt)


df_intensity <- proteingroups_tidyed$intensity_matrix


sparsity <- rowSums(df_intensity > 0) # here sparsity is number of present values

index_all_na_rows <- which(sparsity == 0)

df_intensity <- df_intensity[-index_all_na_rows,,drop = FALSE]

protein_primary_ids <- proteingroups_tidyed$protein_primary_ids[-index_all_na_rows]



```

# Intro

**This report provides some basic description of the proteingroups identificaiton from maxquant database search. **
**The report is based on the proteinGroups.txt, without defined experimental design/grouping information**
**Users can use this to quickly check the overal quality of the experiment**


# Take home figures



*  **Number of contaminant: ** `r proteingroups_tidyed$n_contaminant`

*  **Number of reversed: ** `r proteingroups_tidyed$n_reversed`

*  **Number of qualified proteingroups: ** `r nrow(proteingroups_tidyed$intensity_matrix)`

*  **Number of quailfied proteingroups without intensity information:  ** `r length(index_all_na_rows)`

*  **Number of experiment:  ** `r ncol(df_intensity)`

*  **All experiments:  ** `r colnames(df_intensity)`




# Unique peptide number


```{r echo=FALSE,fig.width= 8,fig.height=6}

peptide_summary <- as.data.frame(table(proteingroups_tidyed$n_unique_peptides))
colnames(peptide_summary) = c("Unique_Peptides", "Freq")


p<- ggplot2::ggplot(data = peptide_summary)+
  geom_col(aes(x = Unique_Peptides,y = Freq))+
  theme_bw()
plotly::ggplotly(p)




```


# Protein Score distribution


```{r echo=FALSE,fig.width= 8,fig.height=6}

df_score <- data.frame(Score  = proteingroups_tidyed$score)


p<- ggplot2::ggplot(data=df_score, aes(Score)) +
  geom_histogram(aes(y=..density..),color= "black",fill="white")+
  geom_density(alpha=.2, fill="#FF6666")+
  theme_bw()

plotly::ggplotly(p)

```


# Intensity distribution

Columns starting with "LFQ_intensity_" will be selected if exist (meaning LFQ option was checked for label free quantification, which is deliberately processed by Maxquant already), otherwise Intensity_ columns will be used instead for protein experession.





```{r echo=FALSE,fig.width= 8,fig.height=6}

df_sparsity <- as.data.frame(table(sparsity))


p<- ggplot2::ggplot(data = df_sparsity)+
  geom_col(aes(x = sparsity,y = Freq))+
  xlab("Presence in the Protein Intensity Matrix")+
  theme_bw()
plotly::ggplotly(p)



```


```{r echo=FALSE,fig.width= 8,fig.height=6}

#note: in this log10 intenisty matrix, 0 is converted into infinity, therefore will not show in the box plot, and does not affect the distribution
df_intensity_log10 <-log10(df_intensity)

data_matrix_log10_melt<-reshape2::melt(as.matrix(df_intensity_log10))

colnames(data_matrix_log10_melt) <- c("Proteins", "Samples", "Log10(Inensity)")

p<-ggplot(data_matrix_log10_melt)+
    geom_boxplot(aes(x = Samples, y = `Log10(Inensity)`)) +
    theme_bw()+
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
  

plotly::ggplotly(p)


```


`r if(ncol(df_intensity) > 1){ "# Overall Expression Profile   ## Heatmap"} `



```{r echo=FALSE,fig.width= 8,fig.height=6}

# only if there is experiment desgin and  number of columns in protein experssion value is more than 1, 
if(ncol(df_intensity) > 1){ 
  df_intensity_log10 <-log10(df_intensity+1)
  rownames(df_intensity_log10) <- protein_primary_ids
  
  d3heatmap(df_intensity_log10,show_grid = FALSE, color = "OrRd")


  
}


```



`r if(ncol(df_intensity) > 1){ "## PCA"}`


```{r echo=FALSE,fig.width= 8,fig.height=6}

if(ncol(df_intensity) > 1){ 
  
  PCA_result <- prcomp(t(df_intensity_log10))
  loading <- as.data.frame(PCA_result$x)
  loading$Sample.Name <- rownames(loading)
  plot_ly(loading, x = ~PC1, y = ~PC2, z = ~PC3) %>%
  add_markers() %>%
  add_text(text = row.names(loading))

}


```

`r if(ncol(df_intensity) > 1){ "**Scree plot shows the performance of the PCA analysis, the more the top procomponets accounts, the better separation**"}`


```{r echo=FALSE,fig.width= 8,fig.height=6}

  # screen plot
if(ncol(df_intensity) > 1){ 

  sd <- PCA_result$sde
  var <- sd^2
  var.percent <- var/sum(var) * 100
  
  PCs <- paste("PC", 1:length(var.percent))
  
  
  df_scree <- data.frame(PC  = factor(PCs, levels = PCs), ratio =var.percent)
  
  
  plot_ly(data = df_scree,x = ~PC, y = ~ratio, type = "bar") %>% 
    layout(title = "Scree Plot of Principle components", xaxis = list(title = 'Principle Component'), yaxis = list(title = 'Variance(%)'))
  

  }


```



# Download the neat table

**The table is clean proteingroups expression table, with reversed and contaminant removed. Only the first protein id was used to represent the proteingroups.**
**The values are the LFQ(with label free quantification turned on) or raw protein intensity from Maxquant output. **
**You can download/export and start from this table for downstream analysis using the (i)Metalab faminly apps. **


```{r echo=FALSE,fig.width= 8,fig.height=6}


df_intensity <- as.data.frame(df_intensity)
rownames(df_intensity) <- protein_primary_ids


DT::datatable(df_intensity, extensions = 'Buttons',options = list(dom = "Blfrtip",scrollX=TRUE, buttons =       list('copy', 'print', list(
        extend = 'collection',
        buttons = c('csv', 'excel', 'pdf'),
        text = 'Download'
      ))))




```



```{r}
# library(ggforce)
# 
# a1 <-100
# a2 <- 500
# a3 <- 1000
# 
# r1 <- sqrt(a1)
# r2 <- sqrt(a2)
# r3 <-sqrt(a3)
# 
# 
# 
# 
# circles <- data.frame(
#   x0 = rep(0,3),
#   y0 = sort(c(r1,r2,r3),decreasing =TRUE),
#   r = sort(c(r1,r2,r3),decreasing =TRUE)
# )
# 
# ggplot(data = circles) +
#   geom_circle( aes(x0 = x0, y0 = y0, r = r, alpha = rev(r)), color = "orange", fill= "orange",show.legend = c("a","b","c") )+
#   coord_fixed()+
#   theme_void()




```





>>>>>>> 1ab705e5efeb58aab665e1d0dd18321bfc231e8b
